---
title: "Introduction to Data Science"
subtitle: "Week 6: Maps and Data Visualization"
author: "Ugur Aytun"
#date: "`r Sys.Date()`"
institute: "METU, Department of Economics | ECON 413"
output:
  xaringan::moon_reader:
    chakra: "C:/Users/Lenovo/Documents/GitHub/uguraytun/docs/libs/remark-latest.min.js"
    css: ["default", "C:/Users/Lenovo/Downloads/ub_theme.css"]
    includes:
      after_body: "C:/Users/Lenovo/Downloads/ub_theme.html"
    self_contained: true
    nature:
      highlightStyle: magula
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```








---
# Maps

--

- Maps are a great way to visualize data that has a geographic component.

--

- They can help us understand spatial relationships and patterns in the data.


--

- In R, we can create maps using the `sf` package for spatial data and the `ggplot2` package for visualization.


--

- The `sf` package provides a simple and consistent way to work with spatial data in R, while `ggplot2` allows us to create beautiful and customizable visualizations.


--

- In this week, we will learn how to create maps using the `sf` and `ggplot2` packages in R.



---
# Loading Required Libraries and set the working directory

```{r}

library(data.table)
library(sf)
library(ggplot2)
library(maps)
library(scales)  # for squish
library(dplyr)

setwd("H:/My Drive/ECON413")

```




















