---
title: "Introduction to Data Science"
subtitle: "Week 6: Maps and Data Visualization"
author: "Ugur Aytun"
#date: "`r Sys.Date()`"
institute: "METU, Department of Economics | ECON 413"
output:
  xaringan::moon_reader:
    chakra: "C:/Users/Lenovo/Documents/GitHub/uguraytun/docs/libs/remark-latest.min.js"
    css: ["default", "C:/Users/Lenovo/Downloads/ub_theme.css"]
    includes:
      after_body: "C:/Users/Lenovo/Downloads/ub_theme.html"
    self_contained: true
    nature:
      highlightStyle: magula
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```







---
# Maps

--

- Maps are a great way to visualize data that has a geographic component.

--

- They can help us understand spatial relationships and patterns in the data.


--

- In R, we can create maps using the `sf` package for spatial data and the `ggplot2` package for visualization.


--

- The `sf` package provides a simple and consistent way to work with spatial data in R, while `ggplot2` allows us to create beautiful and customizable visualizations.


--

- In this week, we will learn how to create maps using the `sf` and `ggplot2` packages in R.




---
# Loading Required Libraries and set the working directory

.scrollable[

```{r}

library(data.table)
library(sf)
library(ggplot2)
library(maps)
library(scales)  # for squish
library(dplyr)

setwd("H:/My Drive/ECON413")

```
]






---
# Load the shape file and check the names of the columns and map

.scrollable[

```{r}
setwd("H:/My Drive/ECON413")

turkey_sf <- read_sf(dsn = "Data/Turkey Administrative Borders", 
                 layer = "TUR_adm1", geometry_column = 1L)


print(names(turkey_sf))
ggplot(data = turkey_sf) + geom_sf()

```
]




---
# Load the shape file and check the names of the columns and map (cont'd)

--

- st_union is used to combine multiple geometries into a single geometry. This helps us to combine the borders of provinces that are split into multiple parts.

--

- The `group_by` function is used to group the data by a specific column, in this case, `NAME_1`, which represents the province names.

.scrollable[

```{r}
setwd("H:/My Drive/ECON413")

turkey_sf$NAME_1[turkey_sf$NAME_1 == "Zinguldak"] <- "Zonguldak"

turkey_sf$NAME_1[turkey_sf$NAME_1 == "Aksaray"] <- "Nigde"
turkey_sf$NAME_1[turkey_sf$NAME_1 == "Karabük"] <- "Zonguldak"
turkey_sf$NAME_1[turkey_sf$NAME_1 == "Bartın"] <- "Zonguldak"
turkey_sf$NAME_1[turkey_sf$NAME_1 == "Bayburt"] <- "Gümüshane"
turkey_sf$NAME_1[turkey_sf$NAME_1 == "Karaman"] <- "Konya"
turkey_sf$NAME_1[turkey_sf$NAME_1 == "Kinkkale"] <- "Ankara"
turkey_sf$NAME_1[turkey_sf$NAME_1 == "Batman"] <- "Siirt"
turkey_sf$NAME_1[turkey_sf$NAME_1 == "Sirnak"] <- "Siirt"
turkey_sf$NAME_1[turkey_sf$NAME_1 == "Ardahan"] <- "Kars"
turkey_sf$NAME_1[turkey_sf$NAME_1 == "Iğdır"] <- "Kars"
turkey_sf$NAME_1[turkey_sf$NAME_1 == "Yalova"] <- "Istanbul"
turkey_sf$NAME_1[turkey_sf$NAME_1 == "Kilis"] <- "Gaziantep"
turkey_sf$NAME_1[turkey_sf$NAME_1 == "Osmaniye"] <- "Adana"
turkey_sf$NAME_1[turkey_sf$NAME_1 == "Düzce"] <- "Bolu"


turkey_67 <- turkey_sf %>%
  group_by(NAME_1) %>%
  summarise(geometry = st_union(geometry), .groups = "drop")



# # old map with 67 provinces
ggplot(data = turkey_67) + geom_sf()

# new map with 81 provinces
ggplot(data = turkey_sf) + geom_sf()



# create license plate numbers
turkey_67$ilkodu <- 0
turkey_67$ilkodu[turkey_67$NAME_1 == "Adana"] <- 1
turkey_67$ilkodu[turkey_67$NAME_1 == "Adiyaman"] <- 2
turkey_67$ilkodu[turkey_67$NAME_1 == "Afyon"] <- 3
turkey_67$ilkodu[turkey_67$NAME_1 == "Agri"] <- 4
turkey_67$ilkodu[turkey_67$NAME_1 == "Amasya"] <- 5
turkey_67$ilkodu[turkey_67$NAME_1 == "Ankara"] <- 6
turkey_67$ilkodu[turkey_67$NAME_1 == "Antalya"] <- 7
turkey_67$ilkodu[turkey_67$NAME_1 == "Artvin"] <- 8
turkey_67$ilkodu[turkey_67$NAME_1 == "Aydin"] <- 9
turkey_67$ilkodu[turkey_67$NAME_1 == "Balikesir"] <- 10
turkey_67$ilkodu[turkey_67$NAME_1 == "Bilecik"] <- 11
turkey_67$ilkodu[turkey_67$NAME_1 == "Bingöl"] <- 12
turkey_67$ilkodu[turkey_67$NAME_1 == "Bitlis"] <- 13
turkey_67$ilkodu[turkey_67$NAME_1 == "Bolu"] <- 14
turkey_67$ilkodu[turkey_67$NAME_1 == "Burdur"] <- 15
turkey_67$ilkodu[turkey_67$NAME_1 == "Bursa"] <- 16
turkey_67$ilkodu[turkey_67$NAME_1 == "Çanakkale"] <- 17
turkey_67$ilkodu[turkey_67$NAME_1 == "Çankiri"] <- 18
turkey_67$ilkodu[turkey_67$NAME_1 == "Çorum"] <- 19
turkey_67$ilkodu[turkey_67$NAME_1 == "Denizli"] <- 20
turkey_67$ilkodu[turkey_67$NAME_1 == "Diyarbakir"] <- 21
turkey_67$ilkodu[turkey_67$NAME_1 == "Edirne"] <- 22
turkey_67$ilkodu[turkey_67$NAME_1 == "Elazığ"] <- 23
turkey_67$ilkodu[turkey_67$NAME_1 == "Erzincan"] <- 24
turkey_67$ilkodu[turkey_67$NAME_1 == "Erzurum"] <- 25
turkey_67$ilkodu[turkey_67$NAME_1 == "Eskisehir"] <- 26
turkey_67$ilkodu[turkey_67$NAME_1 == "Gaziantep"] <- 27
turkey_67$ilkodu[turkey_67$NAME_1 == "Giresun"] <- 28
turkey_67$ilkodu[turkey_67$NAME_1 == "Gümüshane"] <- 29

turkey_67$ilkodu[turkey_67$NAME_1 == "Hakkari"] <- 30
turkey_67$ilkodu[turkey_67$NAME_1 == "Hatay"] <- 31
turkey_67$ilkodu[turkey_67$NAME_1 == "Isparta"] <- 32
turkey_67$ilkodu[turkey_67$NAME_1 == "Mersin"] <- 33
turkey_67$ilkodu[turkey_67$NAME_1 == "Istanbul"] <- 34
turkey_67$ilkodu[turkey_67$NAME_1 == "Izmir"] <- 35
turkey_67$ilkodu[turkey_67$NAME_1 == "Kars"] <- 36
turkey_67$ilkodu[turkey_67$NAME_1 == "Kastamonu"] <- 37
turkey_67$ilkodu[turkey_67$NAME_1 == "Kayseri"] <- 38
turkey_67$ilkodu[turkey_67$NAME_1 == "Kirklareli"] <- 39

turkey_67$ilkodu[turkey_67$NAME_1 == "Kirsehir"] <- 40
turkey_67$ilkodu[turkey_67$NAME_1 == "Kocaeli"] <- 41
turkey_67$ilkodu[turkey_67$NAME_1 == "Konya"] <- 42
turkey_67$ilkodu[turkey_67$NAME_1 == "Kütahya"] <- 43
turkey_67$ilkodu[turkey_67$NAME_1 == "Malatya"] <- 44
turkey_67$ilkodu[turkey_67$NAME_1 == "Manisa"] <- 45
turkey_67$ilkodu[turkey_67$NAME_1 == "K. Maras"] <- 46
turkey_67$ilkodu[turkey_67$NAME_1 == "Mardin"] <- 47
turkey_67$ilkodu[turkey_67$NAME_1 == "Mugla"] <- 48
turkey_67$ilkodu[turkey_67$NAME_1 == "Mus"] <- 49



turkey_67$ilkodu[turkey_67$NAME_1 == "Nevsehir"] <- 50
turkey_67$ilkodu[turkey_67$NAME_1 == "Nigde"] <- 51
turkey_67$ilkodu[turkey_67$NAME_1 == "Ordu"] <- 52
turkey_67$ilkodu[turkey_67$NAME_1 == "Rize"] <- 53
turkey_67$ilkodu[turkey_67$NAME_1 == "Sakarya"] <- 54
turkey_67$ilkodu[turkey_67$NAME_1 == "Samsun"] <- 55
turkey_67$ilkodu[turkey_67$NAME_1 == "Siirt"] <- 56
turkey_67$ilkodu[turkey_67$NAME_1 == "Sinop"] <- 57
turkey_67$ilkodu[turkey_67$NAME_1 == "Sivas"] <- 58
turkey_67$ilkodu[turkey_67$NAME_1 == "Tekirdag"] <- 59


turkey_67$ilkodu[turkey_67$NAME_1 == "Tokat"] <- 60
turkey_67$ilkodu[turkey_67$NAME_1 == "Trabzon"] <- 61
turkey_67$ilkodu[turkey_67$NAME_1 == "Tunceli"] <- 62
turkey_67$ilkodu[turkey_67$NAME_1 == "Sanliurfa"] <- 63
turkey_67$ilkodu[turkey_67$NAME_1 == "Usak"] <- 64
turkey_67$ilkodu[turkey_67$NAME_1 == "Van"] <- 65
turkey_67$ilkodu[turkey_67$NAME_1 == "Yozgat"] <- 66
turkey_67$ilkodu[turkey_67$NAME_1 == "Zonguldak"] <- 67

```
]





---
# Load the output tariff exposure datasets of provinces

.scrollable[
```{r}
setwd("H:/My Drive/ECON413")

output_exposure_province <- setDT(fread("Data/map_output.csv"))
setnames(output_exposure_province, c("il", "exposure"), c("ilkodu", "exposure_output"))

# merge turkey_67 and output_exposure_province
turkey_67 <- merge(turkey_67, 
                   output_exposure_province, 
                   by = c("ilkodu"), 
                   all.x = TRUE)
turkey_67$exposure_output[is.na(turkey_67$exposure_output)] <- 0
                   


# plot
ggplot(data = turkey_67) + 
  geom_sf(aes(fill = exposure_output), color = "white") + 
  geom_sf_text(aes(label = NAME_1) , color = "yellow", size = 2)  + 
  labs(title = "Output exposure across provinces") +
  scale_fill_gradient(
    name = "Exposure Level",
    low = "lightblue", high = "darkblue",
    limits = c(min(turkey_67$exposure_output), 10),  # set upper limit
    oob = squish  # values >5 will be colored as if they were 10
  ) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = c(0.075, 0.20))



```
]







---
# Load the input exposure datasets of provinces

.scrollable[
```{r}
setwd("H:/My Drive/ECON413")

input_exposure_province <- setDT(fread("Data/map_input.csv"))
setnames(input_exposure_province, c("il", "exposure"), c("ilkodu", "exposure_input"))




# merge turkey_67 and output_exposure_province
turkey_67 <- merge(turkey_67, 
                   input_exposure_province, 
                   by = c("ilkodu"), 
                   all.x = TRUE)
turkey_67$exposure_input[is.na(turkey_67$exposure_input)] <- 0




# plot
ggplot(data = turkey_67) + 
  geom_sf(aes(fill = exposure_input), color = "white") + 
  geom_sf_text(aes(label = NAME_1) , color = "yellow", size = 2)  + 
  labs(title = "Input exposure across provinces") +
  scale_fill_gradient(
    name = "Exposure Level",
    low = "lightblue", high = "darkblue",
    limits = c(min(turkey_67$exposure_input), 2),  # set upper limit
    oob = squish  # values >5 will be colored as if they were 2
  ) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = c(0.075, 0.20))

```
]