---
title: "Introduction to Data Science"
subtitle: "Week 8: Regression analysis"
author: "Ugur Aytun"
#date: "`r Sys.Date()`"
institute: "METU, Department of Economics | ECON 413"
output:
  xaringan::moon_reader:
    chakra: "C:/Users/Lenovo/Documents/GitHub/uguraytun/docs/libs/remark-latest.min.js"
    css: ["default", "C:/Users/Lenovo/Downloads/ub_theme.css"]
    includes:
      after_body: "C:/Users/Lenovo/Downloads/ub_theme.html"
    self_contained: true
    nature:
      highlightStyle: magula
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```







---
# Maps

--

- I show you the regression analysis in the next week with a brief introduction to the topic.


--

- Regression analysis is a statistical method used to understand the relationship between variables, often used for prediction and forecasting.

--

- We can also infer causal relationships, but we need to be careful about the assumptions we make.

--

- `fixest` package is used for regression analysis in R, which allows for efficient estimation of linear models with fixed effects.




---
# Key concepts in regression analysis

--

- Outcome variable: The variable we are trying to explain or predict.


--

- Control variable: Additional variables included in the model to account for other factors that may influence the outcome variable.

--

- We can also infer causal relationships, but we need to be careful about the assumptions we make.

--

- Economic theory or intuition: The underlying economic principles or reasoning that guide the choice of variables in the regression model.





---
# Random sample of data

.scrollable[
```{r}

# install.packages("fixest") and set the working directory

#library('R.utils')
library(data.table)
library(sf)
library(ggplot2)
library(maps)
library(scales)  # for squish
library(dplyr)

setwd("H:/My Drive/ECON413")


# install.packages("fixest")

library(fixest)
library(data.table)

# create a random dataset

set.seed(123)
n <- 1000
x1 <- rnorm(n)
x2 <- rnorm(n)
x3 <- rnorm(n)

# create a data frame

df <- data.frame(y = rnorm(n), x1 = x1, x2 = x2, x3 = x3)

# create a data table

df <- as.data.table(df)

# create a data frame with a factor variable

df <- data.frame(y = rnorm(n), x1 = x1, x2 = x2, x3 = x3, group = sample(1:10, n, replace = TRUE))


# regression with fixest

reg1 <-  feols(y # dependent variable
               ~ x1 + x2 + x3 # independent variables
               | group, # fixed effects
               data = df) # data frame
summary(reg1) # summary of the regression 

# interpret the results
# the coefficients of x1, x2, and x3 are the average effect of these variables on y
# the fixed effects are the average effect of the each group (related to reference category) on y

# tabualte the results with more information
table_1 <- etable(reg1) 
table_1


# save as a latex format
etable(reg1, file = "results/table_1.tex") # save as a tex format. 
# What is LaTeX?


# add cluster. What is cluster in the regression analysis?

reg1 <-  feols(y # dependent variable
               ~ x1 + x2 + x3 # independent variables
               | group, # fixed effects
               data = df,
               cluster = c("group")) # data frame
summary(reg1) # summary of the regression 



```
]



---
# Regression analysis with a real dataset
.scrollable[
```{r}
library(WDI)

series <-  WDI_data$series
countries <-  WDI_data$country


# Revisiting the relationship between manufacturing industry and GDP per capita (Kaldor Law #1)

dat2 <- WDI(indicator = c("NY.GDP.MKTP.KD.ZG", # GDP growth
                         "NV.IND.MANF.KD.ZG"), # manufacturing growth
           country = c("all"), start = 1960, end = 2025)


# without fixed effects
reg1 <- feols(NY.GDP.MKTP.KD.ZG ~  # gdp growth
                NV.IND.MANF.KD.ZG, # manufacturing growth
              cluster = c("country"), # cluster
              data = dat2)

summary(reg1)



reg2 <- feols(NY.GDP.MKTP.KD.ZG ~ 
                NV.IND.MANF.KD.ZG |
                country + year, # fixed effects
              cluster = c("country"), # cluster
              data = dat2)

summary(reg2)


etable(reg1, reg2)

```
]






---
# Marginal propenity to consumption (MPC) and the consumption function
.scrollable[
```{r}
# marginal propensity to consume (MPC)
# MPC is the change in consumption divided by the change in income
# MPC = change in consumption / change in income
# lets run a regression

consumption_datasets <- WDIsearch(string = "consumption", 
                                  field = "name", 
                                  short = TRUE, cache = NULL)

gdp_datasets <- WDIsearch(string = "gdp", 
                                  field = "name", 
                                  short = TRUE, cache = NULL)






dat3 <- WDI(indicator = c("NE.CON.TOTL.KD", # Final consumption expenditure (constant 2015 US$)
                          "NY.GDP.MKTP.KD"), # GDP
           country = c("all"), start = 1960, end = 2025)


reg4 <- feols(log(NE.CON.TOTL.KD) ~ 
                log(NY.GDP.MKTP.KD) |
                country + year, # fixed effects
              cluster = c("country"), # cluster
              data = dat3)


summary(reg4)



```
]
