---
title: "Introduction to Data Science"
subtitle: "Week 7: Maps (cont'd)"
author: "Ugur Aytun"
#date: "`r Sys.Date()`"
institute: "METU, Department of Economics | ECON 413"
output:
  xaringan::moon_reader:
    chakra: "C:/Users/Lenovo/Documents/GitHub/uguraytun/docs/libs/remark-latest.min.js"
    css: ["default", "C:/Users/Lenovo/Downloads/ub_theme.css"]
    includes:
      after_body: "C:/Users/Lenovo/Downloads/ub_theme.html"
    self_contained: true
    nature:
      highlightStyle: magula
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```







---
# Maps

--

- This week we proceed with the topic of maps.


--

- Last week we uploaded our own map of Turkey, which was a shapefile containing the administrative borders of Turkey's provinces.


--

- New packages are introduced, such as `rnaturalearth` and `rnaturalearthdata` for geographical data without requiring shapefiles, and `sf` for handling spatial data in R.




---
# Loading Required Libraries, set the working directory and load the map of the world

.scrollable[
```{r}




library(sf)
library("rnaturalearth")
library("rnaturalearthdata")


library("ggplot2")
library(data.table)


world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)




```
]




---
# World map

.scrollable[
```{r}
# plot the world map
ggplot(data = world) + geom_sf()

# plot the world map with a black border
ggplot(data = world) + geom_sf() + theme_bw()

# plot the world map with a black border and no background
ggplot(data = world) + geom_sf() + theme_void()


# plot the world map with a black border and no background
ggplot(data = world) + geom_sf() + theme_bw() +
  labs(title = "World map", 
       subtitle = paste0(length(unique(world$sov_a3)), " countries"), 
       x = "Longitude", y = "Latitude")


# create titles object
ctitles <-   labs(title = "World map", 
                  subtitle = paste0(length(unique(world$sov_a3)), " countries"), 
                  x = "Longitude", y = "Latitude")

# add a title
ggplot(data = world) + 
  geom_sf() + 
  theme_bw() + ctitles


# fill the countries with a color
ggplot(data = world) + 
  geom_sf(color = "black", fill = "lightblue") + 
  theme_bw() + ctitles


# add coordinate numbers
ggplot(data = world) + 
  geom_sf(color = "black", fill = "lightblue") + 
  coord_sf(ylim = c(-65, 90), expand = FALSE) +
  theme_bw() + ctitles


```
]


---
# World and Turkey map

.scrollable[
```{r}

# color the countries by economic development
ggplot(data = world) + 
  geom_sf(aes(fill = economy)) + 
  coord_sf(ylim = c(-65, 90), expand = FALSE) +
  theme_bw() + ctitles +
  scale_fill_viridis_d(option = "plasma")



# color the countries by population
ggplot(data = world) + 
  geom_sf(aes(fill = pop_est)) + 
  coord_sf(ylim = c(-65, 90), expand = FALSE) +
  theme_bw() + ctitles +
  scale_fill_viridis_c(option = "plasma", trans = "sqrt")





# color the countries by population (inverse color scale)
ggplot(data = world) + 
  geom_sf(aes(fill = pop_est)) + 
  coord_sf(ylim = c(-65, 90), expand = FALSE) +
  theme_bw() + ctitles +
  scale_fill_viridis_c(option = "plasma", trans = "sqrt", direction = -1)





# color the countries by population (inverse color scale) with a different projection
ggplot(data = world) + 
  geom_sf(aes(fill = pop_est)) + 
  theme_bw() + ctitles +
  scale_fill_viridis_c(option = "plasma", trans = "sqrt", direction = -1) +
  coord_sf(crs = st_crs(3035))



# Turkey map

turkey <- ne_states(country = "turkey", returnclass = "sf")

class(turkey)





ggplot(data = turkey) + 
  geom_sf() + 
  theme_void() + 
  labs(title = "Turkey map")


# add province name
ggplot(data = turkey) + 
  geom_sf() + 
  geom_sf_text(aes(latitude, longitude, label = name), size = 2) +
  theme_void() + 
  labs(title = "Turkey map")


# different color
ggplot(data = turkey) +
  geom_sf(color = "black", fill = "red") + 
  geom_sf_text(aes(latitude, longitude, label = name), size = 2) +
  theme_void() + 
  labs(title = "Turkey map")






```

]