---
title: "Introduction to Data Science"
subtitle: "Week 5: Data Manipulation (Cont'd) and Visualization"
author: "Ugur Aytun"
#date: "`r Sys.Date()`"
institute: "METU, Department of Economics | ECON 413"
output:
  xaringan::moon_reader:
    chakra: "C:/Users/Lenovo/Documents/GitHub/uguraytun/docs/libs/remark-latest.min.js"
    css: "C:/Users/Lenovo/Downloads/ub_theme.css"
    includes:
      after_body: "C:/Users/Lenovo/Downloads/ub_theme.html"
    self_contained: true
    nature:
      highlightStyle: magula
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```








---
# Import larger data

--

- BACI 2022 trade network data by CEPII:


--

- t: year, i: origin (UN country code), j: destination (UN country code), k: six-digit HS product, v: flow value (thousand USD), q: weight (tonnage)


```{r}
library(data.table)
setwd("H:/My Drive/ECON413")
trade_dataset_2022 = setDT(fread("Data/BACI_HS02_Y2022_V202401b.csv"))
trade_dataset_2022[, .N] # number of observations
summary(trade_dataset_2022)

```




---
# Import larger data (cont'd)

--

- Turkey's total export in 2022?


```{r}
setwd("H:/My Drive/ECON413")
trade_dataset_2022 = setDT(fread("Data/BACI_HS02_Y2022_V202401b.csv"))
trade_dataset_2022[i == 792, sum(v, na.rm = TRUE)]

```



--
- Turkey's total export to Germany in 2022?

```{r}
trade_dataset_2022[i == 792 & j == 276, sum(v, na.rm = TRUE)]

```





---
# Import larger data (cont'd)

--

- Create a dataset showing Turkey's total export by each destination country in descending order


```{r}
partners <- trade_dataset_2022[i == 792, 
                               .(total_export = sum(v, na.rm = TRUE)), 
                               by = c("j")][order(-total_export)]
partners
```




---
# Binding two dataset

--

- 2021 data? Let's call it and merge with 2022.


```{r}
setwd("H:/My Drive/ECON413")
trade_dataset_2021 <- setDT(fread("Data/BACI_HS02_Y2021_V202401b.csv"))
trade_dataset <-  rbind(trade_dataset_2022, # 11,113,911 obs.
                        trade_dataset_2021) # 11,358,186 obs.
trade_dataset[, .N]

```

--

- Turkey's total robot export by year


```{r}
trade_dataset[j == 792 & k == 847950, sum(v, na.rm = TRUE), by = c("t")]

```




---
# Merging country and product names

--

- Tired of checking the country codes online? Lets merge them.




```{r}
library(countrycode)

countries <- codelist
countries <- setDT(countries) # convert countries to data.table
names(countries) # see the variable names and find appropriate columns

```






---
# Merging country names

--

- we need to create two dataset to merge origin (i) and destination (j) columns




```{r}


countries_i = countries[, .(i = un, 
                            i_name = un.name.en)]

countries_j = countries[, .(j = un, 
                            j_name = un.name.en)]



# merge origins
trade_dataset <- merge(trade_dataset,
                       countries_i,
                       by = c("i"),
                       all.x = TRUE)


# merge destinations
trade_dataset <- merge(trade_dataset,
                       countries_j,
                       by = c("j"),
                       all.x = TRUE)

```




---
# Check the merged dataset



```{r}


# merge origins
head(trade_dataset)

# Total exports of Turkey to Russia, sorted by year
trade_dataset[i_name == "Türkiye" & j_name == "Russian Federation", 
              sum(v),
              by = c("t")][order(t)]


```





---
# Merge product descriptions



```{r}


library(concordance)

products <- hs1_desc
products <- setDT(products)

head(products)






```





---
# Merge product descriptions

--

- Product codes are string (character) and consists aggregated groupings such as two or four digit. We need to get rid of them.

```{r}



products[, code := as.integer(code)]
products <- products[code > 9999] # five digit products are agriculture sector. We need keep them.

# rename the code name code
setnames(products, c("code"), c("k"))

# now we merge
trade_dataset <- merge(trade_dataset,
                       products,
                       by = c("k"),
                       all.x = TRUE)



```

--

- Some product descriptions may be not available in product dataset. We keep them by adding the option "all.x = TRUE"




---
# Exercise

--

- What is the total export from Türkiye to Russia containing the word "orange" in product description?


--

- We can use the function "str_detect" from the package "stringr" to check if a string contains another string.


```{r}
library(stringr)
trade_dataset[i_name == "Türkiye" & j_name == "Russian Federation" 
              & str_detect(desc, "orange"),
              sum(v), by = c("t")][order(t)]
```


---
# ggplot2

--

- ggplot2 is a data visualization package. It is very powerful and flexible. It is based on the grammar of graphics.

--

- It is a bit complicated at the beginning but it is worth to learn.


--

- It is a bit slow for large datasets. But it is very flexible and powerful.


--

- more? https://ggplot2.tidyverse.org/



---
# ggplot2 (cont'd)

--

- Example: Let's plot the Total external debt of Turkey 
```{r}
library(ggplot2)
library(CBRT)


# data catagories
catagories <- allCBRTCategories

# data groups
groups <- (allCBRTGroups)

# series
series <- (allCBRTSeries)

```








---
# ggplot2 (cont'd)

--

- Example: Let's plot the Total external debt of Turkey 
```{r}
library(ggplot2)
library(CBRT)


mySeries <- getDataSeries("TP.D1TOP", # External debt, A.Total Payments(Million USD) 
                          CBRTKey = "MQbGGYULHO", # my API
                          freq = 5, # monthly
                          startDate="01-01-2013")


mySeries <- setDT(mySeries)

setnames(mySeries, c("TP.D1TOP"), c("external_debt"))


p <-  ggplot(data = mySeries, aes(x = time, y = external_debt)) + 
  geom_line() + theme_minimal() +
  labs(title = "External debt of Turkey",
       x = "Date",
       y = "External debt (million USD)"
  )


```





---
# ggplot2 (cont'd)

--

- Example: Let's plot the Total external debt of Turkey 
```{r}
print(p)

```
